---
title: "eds-221-day-8"
format: html
editor: visual
---

## Review basics of ggplot

```{r}
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(gapminder)
library(sf)
```

## Read in data

```{r}
lizards <- read_csv(here("data_tidy","lizards.csv"))
```

Do some data wrangling to create a new subset from lizards where you:

1.  Limit the data to only contain western whiptails, colorado checkered, and eastern fence lizards
2.  Add a new column with total length converted to cm
3.  Group by species and site
4.  Return a summary table with mean total length and standard deviation of total length by species and site

```{r}
new_lizards <- lizards %>% 
  filter(common_name %in% c("western whiptail", "colorado checkered", "eastern fence")) %>% 
  mutate(total_length_cm = total_length/10) %>% 
  group_by(common_name, site) %>% 
  summarize(mean_total_length = mean(total_length_cm), sd_total_length = sd(total_length_cm))
```

```{r}
ggplot(lizards, aes(total_length, weight, color = weight)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 100),
                     breaks = seq(from = 0, to = 100, by = 20),
                     expand = c(0, 0)) +
  scale_color_gradientn(colors = c("cornflowerblue", "grey", "lightpink"))
```

```{r}
lizard_counts <- lizards %>% 
  count(common_name)
```

```{r}
ggplot(lizard_counts, aes(y = fct_reorder(common_name, n), x = n)) +
  geom_col() 
```

```{r}
ggplot(gapminder, aes(lifeExp, gdpPercap)) +
  geom_point() +
  annotate("text", x = 40, y = 90000, label = "a label") +
  geom_hline(yintercept = 70000, linetype = "dotted", color = "cornflowerblue") +
  geom_vline(xintercept = mean(gapminder$lifeExp), color = "pink")
```

```{r}
gapminder %>% 
  filter(year == 2002, continent == "Europe") %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(shape = 0, size = 4, color = "cornflowerblue") +
  #geom_text(aes(label = country, color = "lightpink")) + # ggrepel 
  gghighlight(country == "Poland", label_key = country)
```

Create a ggplot chart using data only from 2007, then highlight all countries in Europe, adding their country name as the label.

```{r}
gapminder %>% 
  filter(year == 2007) %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country)) +
  gghighlight(continent == "Europe", label_key = country)
```

```{r}
jornada_veg <- read_sf(here("data_raw","spatial_vegetation","doc.kml")) %>% 
  select(name = Name)

ggplot() +
  geom_sf(data = jornada_veg, aes(fill = name), color = NA) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "palettesForR::Lights") +
  theme(panel.background = element_rect(color = "cornflowerblue", size = 5, fill= "lightpink2"), 
        plot.background = element_rect(fill = "grey"),
        panel.grid.major = element_line(color = "dodgerblue"), 
        axis.text.x = element_text(color = "palegreen3"))
```
