---
title: "eds-221-day-8"
format: html
editor: visual
---

## Review basics of ggplot

```{r}
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
```

## Read in data

```{r}
lizards <- read_csv(here("data_tidy","lizards.csv"))
```

Do some data wrangling to create a new subset from lizards where you:

1.  Limit the data to only contain western whiptails, colorado checkered, and eastern fence lizards
2.  Add a new column with total length converted to cm
3.  Group by species and site
4.  Return a summary table with mean total length and standard deviation of total length by species and site

```{r}
new_lizards <- lizards %>% 
  filter(common_name %in% c("western whiptail", "colorado checkered", "eastern fence")) %>% 
  mutate(total_length_cm = total_length/10) %>% 
  group_by(common_name, site) %>% 
  summarize(mean_total_length = mean(total_length_cm), sd_total_length = sd(total_length_cm))
```

```{r}
ggplot(lizards, aes(total_length, weight, color = weight)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 100),
                     breaks = seq(from = 0, to = 100, by = 20),
                     expand = c(0, 0)) +
  scale_color_gradientn(colors = c("cornflowerblue", "grey", "lightpink"))
```

```{r}
lizard_counts <- lizards %>% 
  count(common_name)
```

```{r}
ggplot(lizard_counts, aes(y = fct_reorder(common_name, n), x = n)) +
  geom_col() 
```
